<center><font size="6"><b>Template-based automatic detection on Spix's disc-winged bat social calls</b></font></center>

### 2022-07-13 scripts/study_case_taeniopygia.Rmd
first commit

```{r manual selection table, eval = FALSE, echo = FALSE}

#Read in manual selections made in Raven

mnl_sel_tbl <- imp_raven(path = raw_data_path, name.from.file = TRUE, ext.case = "lower", all.data = TRUE)

glimpse(mnl_sel_tbl)

mnl_sel_tbl <- mnl_sel_tbl %>% 
  dplyr::rename(
    `start` = "Begin Time (s)",
    `end` = "End Time (s)",
    `selec` = "Selection",
    `selection_notes` = "Annotation"
  ) %>% 
  dplyr::mutate(
    bottom.freq = 0.5,
    top.freq = 10
  ) %>% 
  dplyr::select(sound.files, selec, start, end, selec.file, bottom.freq, top.freq, selection_notes)

glimpse(mnl_sel_tbl)

# Calculate signal to noise ratio (SNR)
mnl_sel_tbl_snr <- sig2noise(mnl_sel_tbl, mar = 0.001, type = 1, in.dB = TRUE, before = FALSE, lim.dB = TRUE, bp = c(unique(mnl_sel_tbl$bottom.freq), unique(mnl_sel_tbl$top.freq)), wl = 10, path = raw_data_path)

glimpse(mnl_sel_tbl_snr)

# Write out the combined manual selections out as a selection table (.csv)
write.csv(mnl_sel_tbl_snr, file.path(preproc_path, "manual_selections_Taeniopygia.csv"), row.names = FALSE)

```{r manual extended selection table, eval = FALSE, echo = FALSE}
#Write out the combined manual selections out as an extended selection table (RDS).

est <- selection_table(mnl_sel_tbl_snr, whole.recs = FALSE, extended = TRUE, mar = 0.01, parallel = cores, confirm.extended = FALSE, path = raw_data_path)

glimpse(est)
class(est)

# Update sound file names, which now have "_number" after the ".wav" extension
wav_nms <- names(attr(est, "wave.objects"))
wav_nms <- paste(gsub(".wav", "", wav_nms), ".wav", sep = "")
head(wav_nms) # looks good

est2 <- rename_est_waves(est, wav_nms)

# Looks good
head(names(attr(est2, "wave.objects")))
glimpse(est2)

# Remove the "old.sound.file.name" column
est2 <- est2[, -grep("old.sound.file.name", names(est2))]
glimpse(est2)

# Add in original sound file name, since each sound.file value is currently a unique name per row of the table
check_df <- attr(est2, "check.res")
glimpse(check_df)
head(check_df$sound.files)

# Are all sampling rates and bit rates the same?
# Yes: 44.1 kHz sampling rate and bit rate of 16
check_df %>%
  distinct(sample.rate, bits)

# Does the order of selections match between the EST and this attribute data frame? If can, can do the column addition below of original sound file name, start and end coordinates, selection duration
all(est2$sound.files == check_df$sound.files)

est2$duration <- check_df$duration
est2$original_sound_files <- check_df$orig.sound.files
est2$original_start <- check_df$orig.start
est2$original_end <- check_df$orig.end

# Looks good
glimpse(est2)
class(est2)
length(attr(est2, "wave.objects"))

saveRDS(est2, file.path(preproc_path, "manual_selections_Taeniopygia_est.RDS"))

```{r automated detection parameters, eval = FALSE, echo = FALSE}

est <- readRDS(file.path(preproc_path, "manual_selections_Taeniopygia_est.RDS"))
# glimpse(est)

mnl_sel_tbl <- read.csv(file.path(preproc_path, "manual_selections_Taeniopygia.csv"))
# glimpse(mnl_sel_tbl)

# Use ohun::feature_reference to get parameters that can help automated detection
# ?ohun::feature_reference

# I get an error when using the EST: "# Error in warbleR::gaps(X = reference, pb = FALSE) : 
  # extended selection tables must be created 'by.song' to be used in song.param()"
# feat_ref <- ohun::feature_reference(est, by.sound.file = FALSE,)

feat_ref <- ohun::feature_reference(mnl_sel_tbl, by.sound.file = FALSE)

# Min and max duration of the manual selections
# Use these for automated detection 
min_dur <- floor(feat_ref[["sel.duration", "min"]]) 
max_dur <- ceiling(feat_ref[["sel.duration", "max"]] )
min_gap <- floor(feat_ref[["gap.duration", "min"]])
bp <- c(feat_ref[["bottom.freq", "min"]] - 0.5, feat_ref[["top.freq", "max"]] + 0.5)

min_dur
max_dur
min_gap
bp

# A single file for testing
wav <- "R39Y212_43533.30691176_3_9_8_31_31.wav"

@@

# create catalogs
catalog(X = high_snr, flim = c(0, 11), nrow = 6, ncol = 3, ovlp = 90, height = 15, width = 20, same.time.scale = FALSE, mar = 0.01, wl = 512, gr = FALSE, spec.mar = 0.4, lab.mar = 0.8, rm.axes = TRUE, by.row = TRUE, box = TRUE, pal = viridis, parallel = 10, collevels = seq(-100, 0, 5), path = data_path)

@@

### Split reference annotations for training and testing detection
```{r test and train, eval = TRUE, echo = TRUE}

train_files <- c(
  "BRN7_43435.27985312_12_1_7_46_25.wav",
  "db1_43427.28100313_11_23_7_48_20.wav",
  "Gry35HP_43455.29800260_12_21_8_16_40.wav"
)
test_files <- setdiff(manual_ref$sound.files, train_files)

@@

### Check subset with highest performance (sorted by f1 score)
```{r, eval = FALSE}
# subset with highest performance
sub_opt_det <- opt_det[order(opt_det$f1.score, decreasing = TRUE), ]

@@

### Run detection on test data
```{r, detect on all data, eval = TRUE, echo = TRUE}

opt_det_all <- energy_detector(
  files = test_files, 
  threshold = best_param$threshold, 
  hop.size = 11.6, 
  smooth = best_param$smooth, 
  hold.time = best_param$hold.time, 
  min.duration = best_param$min.duration, 
  max.duration = best_param$max.duration, 
  bp = c(0.5, 10), 
  path = data_path
)
```

@@

### Evaluate detection performance
```{r, diagnose, eval = TRUE, echo = TRUE}

diagnose_detection(reference = test_ref, detection = opt_det_all, by.sound.file = FALSE)

>>>>

@@

high_snr$bottom.freq <- -1
high_snr$top.freq <- 12

# create catalogs
catalog(X = high_snr, flim = c(0, 11), nrow = 6, ncol = 3, ovlp = 90, height = 15, width = 20, same.time.scale = TRUE, mar = 0.01, wl = 512, gr = FALSE, spec.mar = 0, lab.mar = 0.001, rm.axes = TRUE, by.row = TRUE, box = TRUE, pal = viridis, parallel = 10, collevels = seq(-140, 0, 5), path = data_path, img.prefix = "same.time", highlight = TRUE, alpha = 0.2)

@@

### Split reference annotations for training and testing detection
```{r trains set, eval = FALSE, echo = FALSE}

# train set must be this one
train_files <- c(
  "BRN7_43435.27985312_12_1_7_46_25.wav",
  "db1_43427.28100313_11_23_7_48_20.wav",
  "Gry35HP_43455.29800260_12_21_8_16_40.wav"
)

```{r test and train, eval = TRUE, echo = TRUE}

set.seed(450)    
train_files <- sample(unique(manual_ref$sound.files), 3)

test_files <- setdiff(manual_ref$sound.files, train_files)

@@

### Check subset with highest performance (sorted by f1 score)
```{r, eval = FALSE}

@@

### Run detection on test data
```{r, detect on all data, eval = FALSE, echo = TRUE}

det_test <- energy_detector(
  files = test_files, 
  threshold = best_param$threshold, 
  hop.size = 11.6, 
  smooth = best_param$smooth, 
  hold.time = best_param$hold.time, 
  min.duration = best_param$min.duration, 
  max.duration = best_param$max.duration, 
  bp = c(0.5, 10), 
  path = data_path
)
```

```{r , eval = FALSE, echo = FALSE}

write.csv(det_test, file.path(preproc_path, "detection_test_data.csv"), row.names = FALSE)

@@

### Evaluate detection performance

```{r , eval = TRUE, echo = FALSE}

det_test <- read.csv(file.path(preproc_path, "detection_test_data.csv"))

```{r, diagnose, eval = TRUE, echo = TRUE}

diagnose_detection(reference = test_ref, detection = det_test, by.sound.file = FALSE)

### 2022-07-13 scripts/study_case_thyroptera.Rmd
first commit

data_path <- "./data/raw/thyroptera/"
```

@@

```{r get sample of recordings from flight coordination paper, eval = FALSE, echo = FALSE}
fixed_calls <- read.csv( "./data/processed/manually_fixed_call_detections.csv")
table(fixed_calls)
table(fixed_calls$detection)
length(unique(fixed_calls$sound.files))

@@

manual_ref <- read.csv(file.path(data_path, "manual_annotations_thyroptera.csv"), stringsAsFactors = FALSE)
```

@@

# create catalogs
catalog(X = high_snr, flim = c(10, 45), nrow = 6, ncol = 5, ovlp = 90, height = 15, width = 20, same.time.scale = TRUE, mar = 0.005, wl = 512, gr = FALSE, spec.mar = 0.4, lab.mar = 0.8, rm.axes = TRUE, by.row = TRUE, box = TRUE, pal = viridis, parallel = 10, collevels = seq(-100, 0, 5))

@@

train_files <- c(
 "ch1MPI2020-01-17_08-44-19_0000028.wav",
 "ch1MPI2020-01-23_10-30-35_0000106.wav",
 "ch1MPI_2020-01-30_15-06-38_0000193.wav",
 "ch1MPI2020-01-16_11-46-10_0000018.wav",
 "ch1MPI2020-01-16_13-40-58_0000019.wav",
 "ch1MPI2020-01-24_11-28-15_0000124.wav"
)
test_files <- setdiff(manual_ref$sound.files, train_files)

@@

### Get diverse set of templates
```{r cross-correlation detection, eval = FALSE, echo = TRUE}

# find templates
templates <- get_templates(train_ref, path = data_path,  bp = c(10, 50), fast = TRUE, ovlp = , wl = 200, n.templates = 4)

# create ext. selection table
templates_est <- selection_table(templates, extended = TRUE, confirm.extended = FALSE, path = data_path)

templates_est <- rename_est_waves(templates_est, new.sound.files = templates$distance.quantile)

@@

corr_templ <- template_correlator(templates = templates_est,
    path = data_path, files = unique(train_ref$sound.files), parallel = paral, hop.size = 10, ovlp = 70)

@@

opt_detec_train <- optimize_template_detector(
  reference = train_ref, 
  template.correlations = corr_templ_train,
  threshold = seq(0.05, 0.5, 0.01)
  )

@@

```{r, eval = FALSE, echo = FALSE}
templates_est <- readRDS("./data/processed/templates_thyroptera_est.RDS")

@@

opt_detec_train <- optimize_template_detector(
  reference = train_ref, 
  template.correlations = corr_templ_train,
  threshold = seq(0.05, 0.5, 0.01), 
  parallel = paral
  )

opt_detec_train$templates <- gsub("-1", "", opt_detec_train$templates)

@@

### Create recall-precision plot
```{r, eval = FALSE}

# summarize across sound files
diag_train <- summarize_diagnostic(opt_detec_train, time.diagnostics = FALSE)
agg_sd <- aggregate(cbind(recall, precision) ~ threshold + templates, data = opt_detec_train,
    mean)

# plot
ggplot(agg_sd, aes(x = recall, y = precision, group = templates, color = templates)) +
    geom_line() + geom_point() + scale_color_viridis_d(end = 1) + theme_classic()
```

@@

diag_train <- summarize_diagnostic(opt_detec_train, time.diagnostics = FALSE)

# names(diag_train)

agg_sd <- aggregate(cbind(recall, precision) ~ threshold + templates, data = diag_train,
    mean)

ggplot(agg_sd, aes(x = recall, y = precision, group = templates, color = templates)) +
    geom_line() + geom_point() + scale_color_viridis_d(end = 1) + theme_classic()

@@

sub_diag_train <-diag_train[order(diag_train$f1.score, decreasing = TRUE), ]

sub_diag_train <- sub_diag_train[1:10, c("templates","threshold", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]

sub_diag_train
```

@@

# subset with highest performance
sub_diag_train <-diag_train[order(diag_train$f1.score, decreasing = TRUE), ]

sub_diag_train <- sub_diag_train[1:10, c("templates","threshold", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]

# print dynamic table
oa_DT <- datatable(sub_diag_train, editable = list(target = "row"), rownames = FALSE, style = "bootstrap",
    filter = "top", options = list(pageLength = 100, autoWidth = TRUE, dom = "ft"),
    autoHideNavigation = TRUE, escape = FALSE)

formatRound(table = oa_DT, columns = sapply(sub_diag_train, is.numeric), 3)
```

@@

### Run detection on test files
```{r, detect on all data, eval = FALSE, echo = TRUE}

# get correlation vectors for test files
corr_templ_test <- template_correlator(
  templates = templates_est[templates_est$sound.file == "50%", ],
    path = data_path, files = unique(test_ref$sound.files), 
  parallel = paral, 
  hop.size = 10, 
  ovlp = 70
  )

# extract best template-threshold combination
opt_50 <- optimize_fourier_detec[optimize_fourier_detec$templates == "50%", ]
opt_thresh <- opt_50$threshold[which.max(opt_50$f1.score)]

# detect on test files
detec_test <- template_detector(
  template.correlations = corr_templ_test,
    threshold = opt_thresh
  )
```

@@

### Evaluate detection performance on test files
```{r, diagnose detect on all data, eval = FALSE, echo = TRUE}

@@

```{r, eval = FALSE, echo = FALSE}
saveRDS(detec_test, "./data/processed/detection_thyroptera_all.RDS")
```

@@

detec_test <- readRDS("./data/processed/detection_thyroptera_all.RDS")

diagnose_detection(reference = test_ref, detection = detec_test)
```

@@

- Good performance on detecting Spix's disc-winged social calls: F1 score was 0.99 for the training data set and 0.94 for the testing data

>>>>

data_path <- "./data/raw/thyroptera/"

source("~/Dropbox/R_package_testing/ohun/R/get_templates.R")
```

@@

```{r get sample of recordings from flight coordination paper, eval = FALSE, echo = FALSE}

fixed_calls <- read.csv( "./data/processed/manually_fixed_call_detections.csv")
table(fixed_calls)
table(fixed_calls$detection)
length(unique(fixed_calls$sound.files))

@@

manual_ref <- read.csv(file.path(data_path, "manual_annotations_thyroptera.csv"))

@@

# create catalogs
catalog(X = high_snr, flim = c(10, 45), nrow = 6, ncol = 5, ovlp = 90, height = 10, width = 20, same.time.scale = TRUE, mar = 0.005, wl = 512, gr = FALSE, spec.mar = 0, lab.mar = 0.001, rm.axes = TRUE, by.row = TRUE, box = FALSE, pal = viridis, collevels = seq(-100, 0, 5), labels = NULL)

```{r fig for paper, eval = FALSE, echo = FALSE}

catalog(X = high_snr[1:15, ], flim = c(10, 45), nrow = 3, ncol = 5, ovlp = 90, height = 7.5, width = 20, same.time.scale = TRUE, mar = 0.005, wl = 512, gr = FALSE, spec.mar = 0.4, lab.mar = 0.2, rm.axes = TRUE, by.row = TRUE, box = TRUE, pal = viridis, parallel = 1, collevels = seq(-100, 0, 5), img.prefix = "thyroptera_15_calls", labels = NULL)

@@

set.seed(1)

train_files <- sample(unique(manual_ref$sound.files), size = 5)

test_files <- setdiff(manual_ref$sound.files, train_files)

@@

### Get diverse set of templates
```{r get templates, eval = TRUE, echo = TRUE, fig.height=7}
# find templates
templates <- get_templates(train_ref, path = data_path,  bp = c(10, 50), fast = TRUE, ovlp = , wl = 200, n.sub.spaces = 3)
```

```{r get templates 2, eval = FALSE, echo = TRUE}
# create ext. selection table
templates_est <- selection_table(templates, extended = TRUE, confirm.extended = FALSE, path = data_path)

templates_est <- rename_est_waves(templates_est, new.sound.files = templates$template)

@@

@@

opt_detec_train <- optimize_template_detector(
  reference = train_ref, 
  template.correlations = corr_templ_train,
  threshold = seq(0.1, 0.9, 0.05)
  )

@@

templates_est <- readRDS("./data/processed/templates_thyroptera_est.RDS")

@@

opt_detec_train <- optimize_template_detector(
  reference = train_ref, 
  template.correlations = corr_templ_train,
  threshold = seq(0.1, 0.9, 0.05), 
  parallel = paral
  )

opt_detec_train$templates <- gsub("-1$", "", opt_detec_train$templates)

@@

### Create plot with threshold vs f1 score
```{r, eval = TRUE, echo = FALSE}

@@

agg_f1 <- aggregate(f1.score ~ threshold + templates, data = opt_detec_train,
mean)

ggplot(agg_f1, aes(x = threshold, y = f1.score, group = templates, color = templates)) +
geom_line() + geom_point() + scale_color_viridis_d(end = 1) + theme_classic()

@@

opt_detec_train <- opt_detec_train[order(opt_detec_train$f1.score, decreasing = TRUE), ]

sub_opt_detec_train <- opt_detec_train[1:10, c("templates","threshold", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]

sub_opt_detec_train
```

@@

# subset with highest performance
opt_detec_train <- opt_detec_train[order(opt_detec_train$f1.score, decreasing = TRUE), ]

sub_opt_detec_train <- opt_detec_train[1:10, c("templates","threshold", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]

# print dynamic table
oa_DT <- datatable(sub_opt_detec_train, editable = list(target = "row"), rownames = FALSE, style = "bootstrap",
    filter = "top", options = list(pageLength = 100, autoWidth = TRUE, dom = "ft"),
    autoHideNavigation = TRUE, escape = FALSE)

formatRound(table = oa_DT, columns = sapply(sub_opt_detec_train, is.numeric), 3)

@@

### Run detection on test files

We used a threshold of 0.45 as te centroid template seems to behave well across thresholds.
```{r, detect on all data 67, eval = FALSE, echo = TRUE}

# get correlation vectors for test files
corr_templ_test <- template_correlator(
  templates = templates_est[templates_est$sound.file == "centroid", ],
    path = data_path, files = unique(test_ref$sound.files), 
  hop.size = 10, 
  ovlp = 70
  )

# detect on test files
detec_test <- template_detector(
  template.correlations = corr_templ_test,
    threshold = 0.45
  )

@@

### Evaluate detection performance on test files
```{r, diagnose detect on all data 67, eval = FALSE, echo = TRUE}

@@

saveRDS(detec_test, "./data/processed/detection_thyroptera_all_67%_template.RDS")

@@

detec_test <- readRDS("./data/processed/detection_thyroptera_all_67%_template.RDS")

dd <-diagnose_detection(reference = test_ref, detection = detec_test)

dd

dd$f1.score

@@

- Good performance on detecting Spix's disc-winged social calls: F1 score was `r round(dd$f1.score, 3)` for the training data set and `r round(max(opt_detec_train$f1.score), 3)` for the testing data

- Some merged positives were found but notice that there are some overlapping signals in the reference annotation as well (can be checked with `warbleR::overlapping_sels(manual_ref)`)
